---
title: "DATA 607 - Project 4"
author: "Jered Ataky, Magnus Skonberg"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: architect
    highlight: github
    df_print: paged
    smooth_scroll: yes
    toc_collapsed: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
    toc_float: yes
    code_folding: "hide"
---

```{r echo = FALSE}
knitr::opts_chunk$set(eval = TRUE, results = TRUE, fig.show = "asis", message = FALSE)
```


```{r load-packages, include=FALSE}
library(tidyverse)
library(knitr)
library(R.utils)
library(DT)
library(tm)
library(wordcloud)
library(data.table)
library(e1071)
```

### Background

The focus of this project is **document classification**. 

For this project, we will start with a corpus dataset, unzip our data, generate a training model that we'll then use to predict the class of new documents (those withheld from the training set or taken from another source), and then analyze the accuracy of our predictive classifier.

### Download Data

We lean on the R.utils library to *automatically* download, bunzip, extract the contents of tar archive into our "emails" directory, and then create a corresponding list of file names from the spam and ham emails available on [spamassassin](https://spamassassin.apache.org/old/publiccorpus/):

```{r}
#Download, bunzip, and untar spam_2 files into "emails" directory
#download.file("http://spamassassin.apache.org/old/publiccorpus/20050311_spam_2.tar.bz2", destfile= "20050311_spam_2.tar.bz2")
#bunzip2("20050311_spam_2.tar.bz2")
#untar("20050311_spam_2.tar", exdir="emails")

#Create corresponding list of file names for spam_2 and exclude cmds file
if (file.exists("emails\\spam_2\\cmds")) file.remove("emails\\spam_2\\cmds")
spam_list = list.files("emails\\spam_2\\") 

#Download, bunzip, and untar easy_ham files into "emails" directory
#download.file("http://spamassassin.apache.org/old/publiccorpus/20030228_easy_ham.tar.bz2", destfile="20030228_easy_ham.tar.bz2")
#bunzip2("20030228_easy_ham.tar.bz2")
#untar("20030228_easy_ham.tar", exdir = "emails")

#Create corresponding list of file names for easy_ham and exclude cmds file
if (file.exists("emails\\easy_ham\\cmds")) file.remove("emails\\easy_ham\\cmds")
ham_list = list.files("emails\\easy_ham\\")
```

As can be noted above, we remove the cmds files prior to creating our corresponding lists and we create our corresponding lists using the list.files() function to produce a character vector of the names of files or directories in the named directory.

*Note: if it's your first time running the code please UNCOMMENT the download.file(), bunzip2(), and untar() portions of the code. Otherwise, comment them out to avoid "file already exists" error messages.*

Check the length of corresponding spam and ham lists:

```{r}
length(spam_list)
length(ham_list)
```

The ham_list contains 2500 emails (64.17% of total) and the spam_list contains 1396 emails (35.85% of total) for a total of 3896 emails that we'll be processing. These values are worth noting as we proceed through the code that follows and especially when we consider naive Bayes later.

Once we have lists of spam and ham emaiils, we set out to build a dataframe of all emails df_mails and 

```{r}
# Build data frame
df_mails <- tibble()
df_mails_folders <- c("emails\\spam_2\\", "emails\\easy_ham")
df_mails_types <- c("spam", "ham")

#Extract type (spam vs. ham) and message of corresponding file to populate df_mails 
for (i in 1: length(df_mails_folders))
  
  {
        type <- df_mails_types[i] #spam or ham
        
        #access files
        l <- tibble(file = dir(df_mails_folders[i],  full.names = TRUE)) %>% 
          #read in email messages
          mutate(messages = map(file, read_lines)) %>%
                #use file name as id, type as spam / ham, and message as content
                transmute(id = basename(file), type = type, messages) %>%
                unnest(messages) #make each element of messages its own row
                df_mails<- bind_rows(df_mails, l)
 }
```

Once we've built out our data frame, we notice that it's HUGE. df_mails contains 389362 observations which was a real pain to process later on so we merged messages based on shared id's (if they came from the same email file):

```{r}
#Merge messages based on shared ids:
new_df <- df_mails[!duplicated(df_mails$id), ]
new_df[, 'messages'] <- aggregate (messages~id, data = df_mails, toString) [,2]
head(new_df)

#Subset data frame to only contain type (ham / spam) and message (email content): 
df_final <- new_df %>%
  select(type, messages)
dim(df_final)

```

From the above output, we note that our dataframe has proper dimensions: 2 columns (type, message) and 3896 rows (email contents).

Now that we have a data frame with exclusively type and message, we can convert our data frame into a corpus, clean the contents of this corpus and create a document term matrix so that later we can visualize the highest frequency words, train a naive Bayes model and observe the test results of this model ...

### Create and Clean Corpus

We start by creating a corpus of all of our email messages and then we clean the corpus by removing punctuation, replacing non-word characters with " ", removing URLs, removing numbes, removing non-content words, removing white space, and homogenizing case for all lower case:

```{r}
# Create corpus from all messages of df_final data frame:
##Should we randomize here ...
text_corpus <- Corpus(VectorSource(df_final$messages))
print(text_corpus)

#Clean corpus rd 1:
clean_corpus <- text_corpus %>%
  tm_map(removePunctuation) %>%
  tm_map(content_transformer(gsub), pattern="\\W",replace=" ")

#Remove URLs:
removeURL <- function(x) gsub("http^\\s\\s*", "", x)%>% 
  clean_corpus <- tm_map(clean_corpus, content_transformer(removeURL))

#Clean corpus rd 2:
clean_corpus <- clean_corpus %>%
  tm_map(removeNumbers) %>%
  tm_map(removeWords, stopwords()) %>% 
  tm_map(stripWhitespace) %>%
  tm_map(tolower)

```
Again, we note the proper amount of documents for text_corpus :)

As a next step, we can create our document term matrix, which is a way of representing the words in clean_corpus as a table (with rows representative of text responses to be analyzed and columns representative of words from the text used in the analysis):

```{r}
#Create DTM using function:
dtm <- DocumentTermMatrix(clean_corpus)
dtm #inspect
```

Note the 100% sparsity. Taking this into account, we then filter our DTM (document term matrix) to reduce sparsity by looking for terms that are in at least ~2.25% of the entries before inspecting the output:

```{r}
#Reduce sparsity 
dtm_filtered <- dtm %>% 
  removeSparseTerms(1-10/length(clean_corpus)) #length(clean_corpus) = 3896
inspect(dtm_filtered)

```
We now have ~98% sparsity and account for 6512 terms rather than more than 95k in the unfiltered DTM.

From this point, we can revisit clean_corpus via visualization. We can make use of the wordcloud() function to visualize the most common 50 words with a minimum count of 1000. Those words are:

```{r}
#Visualize as a wordcloud
wordcloud(clean_corpus, max.words = 50, random.order = FALSE, min.freq=1000)

```
We can note the prominence of "received" as well as some other words and some other jumbled words. Maybe the merging of words and such was a result of the order that we cleaned our corpus (ie. removed white space) ...

At this point we''re ready to enter the final phase. We're ready to ...


### Train and Test Data

We're going to apply the naive Bayes classifier.

To steal from [UC Business Analytics' site](https://uc-r.github.io/naive_bayes):

*Although it is often outperformed by other techniques, and despite the naïve design and oversimplified assumptions, this classifier can perform well in many complex real-world problems. And since it is a resource efficient algorithm that is fast and scales well, it is definitely a machine learning algorithm to have in your toolkit.*

First, we divide the corpus into training and test data. We do so for our "raw" emails (from our data frame df_final), data 

```{r}
#Determine 80% of row number:
##nrow(df_final) #3896
##round(0.8 * 3896,0) #3117

email_raw_train <- df_final[1:3117,]
email_raw_test <- df_final[3118:3896,]

#for whatever reason text_final only sends along 6 elements ...
email_dtm_train <- dtm_filtered[1:3117,] 
email_dtm_test <- dtm_filtered[3118:3896,]

#Since corpus is stored as documents:
email_corpus_train <- clean_corpus[1:3117]
email_corpus_test <- clean_corpus[3118:3896]

```

Once our data's been divided, we subset our training data based on type (spam vs. ham), identify words that appear at least five times, make use of a simple function convert_count() to detect the presence or absence of each word in a message, and then we apply this function to our training and test data:

```{r}

#Separate training data into spam and ham subsets
spam <- subset(email_raw_train, type == "spam")
ham <- subset(email_raw_train, type == "ham")

#Identify words that appear at least 5 times:
five_times_words <- findFreqTerms(email_dtm_train, 5)
length(five_times_words) #how many words are there?
five_times_words[1:5]

#Create DTMs using frequent words:
email_train <- DocumentTermMatrix(email_corpus_train, control=list(dictionary = five_times_words))
email_test <- DocumentTermMatrix(email_corpus_test, control=list(dictionary = five_times_words))

#Convert count info to "Yes" or "No"
convert_count <- function(x) {
  y <- ifelse(x > 0, 1, 0)
  y <- factor(y, levels=c(0,1), labels=c("No", "Yes"))
  y
}

#Convert document-term matrices:
email_train <- apply(email_train, 2, convert_count) 
email_test <- apply(email_test, 2, convert_count)

```

From the outputs above, we see that 6210 words appeared at least 5 times and we can see the first five words in our corresponding list.

At this point we're ready to apply and evaluate the performance of naive Bayes. First we train the model with our training data (email_train), we verify the class of the assigned variable (email_classifier) and then we evaluate the performance of our model with the associated test data (email_test). The result can be seen below:

```{r}
#Create naive Bayes classifier object
email_classifier <- naiveBayes(email_train, factor(email_raw_train$type))
class(email_classifier)

#Evaluate performance on test data
email_test_pred <- predict(email_classifier, newdata=email_test)
table(email_test_pred, factor(email_raw_test$type))

```

Our test set contained 779 elements. Of which the proportion of input was 35.83% spam, 64.17% ham. Thus our prediction far over-estimates the presence of ham messages at 80.23% and under-estimates the presence of spam messages at 19.77% ...

I believe it might be because we didn't randomize spam and ham inputs.


### References

In completing Project 4, we found the following resources useful and applicable:

1. Notre Dame. (2014). **Text mining example: spam filtering** [slide set]. Retrieved from https://www3.nd.edu/~steve/computing_with_data/20_text_mining/text_mining_example.html#/

2. Gorakala, Suresh Kumar. (2013). **Document classification using R** [document]. Retrieved from https://www.r-bloggers.com/2013/07/document-classification-using-r/

3. Kharshit. (2017). **Email spam filtering: Text analysis in R** [sample code]. Retrieved from https://kharshit.github.io/blog/2017/08/25/email-spam-filtering-text-analysis-in-r
